---
title: "Homework 3 Solutions"
author: "Kailey Rishovd"
date: "10/10/2020"
output: github_document
---

```{r setup, include= FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = .6, 
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot.continuous.colour = "viridis", 
  ggplot.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1 

```{r}
data("instacart")

instacart
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 

The dataset is extensive, providing details on the online grocery service. Observations are the level of items in an order by user. There are user / order variables -- user ID, order ID, order day, order hour, and reordered. There are also item variables -- name, aisle, department, and some numeric codes. Structurally, we can see that products are in aisles, aisles are in departments, and the tracker supplies information about each user, the products they buy, and where those products are found around local stores. Users are also tracked for how often they make orders, what they order and reorder, the order in which they shop, and when they shop. In this dataset, we are given the information for over 3 million orders from over 200k users.

#### A little more context...  

**How many aisles are there, and which are most items from?**

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

There are 134 aisles and most items are from the fresh vegetables aisle. 

**Plot that shows the number of items ordered in each aisle...**

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1))
```

From this plot we can see how many items were ordered in each aisle... arranged by aisle with the least to the most items ordered. We can tell things such as, butter is ordered the least and fresh vegetables and fruits are ordered significantly more than other items. 


**Table showing three most popular items in specified aisles... and how many times they have been ordered...**

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(
    rank = min_rank(desc(n))
  ) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

From this table, we can see the 3 most popular items ordered from the "baking ingredients", "dog food care" and "packaged vegetables fruits" aisles - including how many times they have been ordered. The top ranking items in each aisle are LIght Brown Sugar, Snack Sticks Chicken and Rice Recipe Dog Treats, and Organic Baby Spinach, respectively. 

**Table showing the mean hour of the day at which specified items are ordered on each day of the week...**

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour
  ) %>% 
  knitr::kable()
```

In this table, we see the mean hour of the day, per day of the week, that Pink Lady Apples and Coffee Ice Cream are ordered. Coffee Ice Cream is mainly ordered between hour 12 and 16. Pink Lady Apples are ordered around hours 11 and 15. 

## Problem 2 

Load and tidy the data... 

```{r}
accel_df = read_csv("./data/accel_data.csv")

accel_df = 
  accel_df %>%
    pivot_longer(
      activity.1:activity.1440, 
      names_to = "minute_of_day",
      values_to = "activity_count", 
      names_prefix = "activity."
    ) %>%
    mutate(
      Wkend_vs_Wkday = case_when(
        day == "Monday" ~ "Weekday",
        day == "Tuesday" ~ "Weekday",
        day == "Wednesday" ~ "Weekday", 
        day == "Thursday" ~ "Weekday", 
        day == "Friday" ~ "Weekday", 
        day == "Saturday" ~ "Weekend", 
        day == "Sunday" ~ "Weekend",
      )) %>% 
    mutate(
      day = as.factor(day), 
      minute_of_day = as.double(minute_of_day), 
      Wkend_vs_Wkday = as.factor(Wkend_vs_Wkday),
      )
```

After tidying up the data, we are left with a dataset that has `r nrow(accel_df)` rows and `r ncol(accel_df)` columns. It consists of the information for five weeks of accelerometer data for a 63-year-old male with BMI 25 and congestive heart failure. The dataset has variables for each week, day of the week, minute of the day, and activity count at each minute of the day... with additional grouping information like weekend vs. weekday. 

Now we can find total activity for each day and look for trends... 

```{r}
accel_df %>% 
  mutate(
    day = forcats::fct_relevel(day, "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
  ) %>% 
  group_by(day, week) %>% 
  summarize(total_day_activity = sum(activity_count)) %>%
  pivot_wider(
    names_from = day,
    values_from = total_day_activity,
  ) %>% 
  knitr::kable()
```

From the table, we can see a few trends. A few: most activity occurs in the weekdays (Monday-Friday), high weekend activity dropped off after week 3, the two lowest activity days were on the last two Saturdays of observation, weeks two and three have the highest and most consistent activity, and activity counts are most consistently the same on Wednesdays. 

We can also make a plot... 

```{r}
accel_df %>% 
    mutate(
    day = forcats::fct_relevel(day, "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
  ) %>% 
  ggplot(aes(x = minute_of_day, y = activity_count, color = day)) +
  geom_point() +
  geom_line()
```

From the plot, we can see other trends. For example, there is low activity in the beginning of the day and the end of the day (likely due to sleep), there is higher activity in the morning/early afternoon or later in the evening (with a typical drop in the middle of the day), and even more specifically, Sunday activity is generally in the early afternoon and Friday activity is usually later in the evening. 

## Probelm 3 

First, load NY NOAA data... 

```{r}
data("ny_noaa")

ny_noaa
```

The NY NOAA dataset is from the NOAA National Climatic Data Center. In the dataset there are `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. It is meant to provide the public with weather data such as the Global Historical Climatology Network-Daily of weather stations in 180 countries. Variables include the station id, date of record, minimum and maximum temperatures, daily total precipitation, and snowfall and depth. However, the dataset is full of missing values because of stations not reporting or reporting partially... data descriptions report that  about half of the stations only report precipitation. 

**Some more context... **

Let's clean the data a bit... 

```{r}
ny_noaa = 
  ny_noaa %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate_at(vars(year:day), as.integer)
```

